---
title: "Becca Analysis 3/6"
author: "BB"
date: "3/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Notes: cor.test was not performed since the dataset utilized here is the US Census.

w/i the groups, is there correlation
b/w the groups, is there correlation
Freedom Score:
Overall Freedom
Economic Freedom
Personal Freedom
Regulatory Policy
Fiscal Policy

Work Variation:
Professional
Service
Office
Construction
Unemployed
Production

Ethnic Variation:
Native
White
Black
Hispanic
Asian

Measuring income/capita

```{r Functions}
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(pkg,dep=T,repos="http://cran.us.r-project.org"); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)
```

```{r DF Setup}
str(full_2015_varOfInterest)
full_2015_cor <- cor(full_2015_varOfInterest, use = "complete.obs")
```

```{r Cor of Dataset}
loadPkg("corrplot")
corrplot(full_2015_cor)
```

```{r Subsetting}
Freedom_2015 <- subset(full_2015_varOfInterest, select = c(IncomePerCap, OverallFreedom, EconomicFreedom, PersonalFreedom, RegulatoryPolicy, FiscalPolicy))
str(Freedom_2015)

Work_2015 <- subset(full_2015_varOfInterest, select = c(IncomePerCap, Professional, Service, Office, Construction, Production, SelfEmployed, Unemployment))
str(Work_2015)

Ethnic_2015 <- subset(full_2015_varOfInterest, select = c(IncomePerCap, Hispanic, White, Black, Native, Asian))
str(Ethnic_2015)
```


```{r Cor w/i Freedom}
loadPkg("corrplot")
Freedom_2015_cor <- cor(Freedom_2015, use = "complete.obs")
corrplot(Freedom_2015_cor)
corrplot.mixed(Freedom_2015_cor, tl.pos = "lt")
```


```{r Cor w/i Work}
Work_2015_cor <- cor(Work_2015, use = "complete.obs")
corrplot(Work_2015_cor)
corrplot.mixed(Work_2015_cor, tl.pos = "lt")
```


```{r Cor w/i Ethnic Background}
Ethnic_2015_cor <- cor(Ethnic_2015, use = "complete.obs")
corrplot(Ethnic_2015_cor)
corrplot.mixed(Ethnic_2015_cor, tl.pos = "lt")
```


```{r lattice}
#This part is commented out because while it's kind of interesting, it takes a long time for each lattice (since it's looking at 70k obs)
#loadPkg("lattice")
#pairs(Ethnic_2015)
#pairs(Work_2015)
#pairs(Freedom_2015)
```

```{r vif }
#loadPkg("car")
#lm2015 <- lm(IncomePerCap ~ ., data = full_2015_varOfInterest)
#lm2015

#vif(lm2015)
#the vif will return an error statement.  I believe it is because of the NA values for the coefficients in the lm
```




