---
title: "Becca Analysis 3/6"
author: "BB"
date: "3/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

w/i the groups, is there correlation
b/w the groups, is there correlation
Freedom Score:
Overall Freedom
Economic Freedom
Personal Freedom
Regulatory Policy
Fiscal Policy

Work Variation:
Professional
Service
Office
Construction
Unemployed
Production

Ethnic Variation:
Native
White
Black
Hispanic
Asian

Measuring income/capita

```{r Functions}
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(pkg,dep=T,repos="http://cran.us.r-project.org"); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)
```

```{r Cor w/i Freedom}
loadPkg("corrplot")

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
