---
title: "Group Project"
author: "Dinesh Padmanabhan, Jake Lieberfarb,
Luis Ahumada, Noah Feldman, Becca Blacker
"
date: "3/09/2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = F, results = T, message = T, echo = F)
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```


```{r Functions, include=FALSE} 
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(pkg,dep=T,repos="http://cran.us.r-project.org"); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)
```

```{r Remove Outliers, include=FALSE} 
# Fix outliers

outlierKD2 <- function(df, var, rm=FALSE) { 
    #' Original outlierKD functino by By Klodian Dhana,
    #' https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/
    #' Modified to have third argument for removing outliers inwtead of interactive prompt, 
    #' and after removing outlier, original df will not be changed. The function returns the new df, 
    #' which can be saved as original df name if desired.
    #' Check outliers, and option to remove them, save as a new dataframe. 
    #' @param df The dataframe.
    #' @param var The variable in the dataframe to be checked for outliers
    #' @param rm Boolean. Whether to remove outliers or not.
    #' @return The dataframe with outliers replaced by NA if rm==TRUE, or df if nothing changed
    #' @examples
    #' outlierKD2(mydf, height, FALSE)
    #' mydf = outlierKD2(mydf, height, TRUE)
    #' mydfnew = outlierKD2(mydf, height, TRUE)
    dt = df # duplicate the dataframe for potential alteration
    var_name <- eval(substitute(var),eval(dt))
    na1 <- sum(is.na(var_name))
    m1 <- mean(var_name, na.rm = T)
    #par(mfrow=c(2, 2), oma=c(0,0,3,0))
    boxplot(var_name, main="With outliers")
    hist(var_name, main="With outliers", xlab=NA, ylab=NA)
    outlier <- boxplot.stats(var_name)$out
    mo <- mean(outlier)
    var_name <- ifelse(var_name %in% outlier, NA, var_name)
    boxplot(var_name, main="Without outliers")
    hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
    title("Outlier Check", outer=TRUE)
    na2 <- sum(is.na(var_name))
    cat("Outliers identified:", na2 - na1, "\n")
    cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "\n")
    cat("Mean of the outliers:", round(mo, 2), "\n")
    m2 <- mean(var_name, na.rm = T)
    cat("Mean without removing outliers:", round(m1, 2), "\n")
    cat("Mean if we remove outliers:", round(m2, 2), "\n")
    
    # response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
    # if(response == "y" | response == "yes"){
    if(rm){
        dt[as.character(substitute(var))] <- invisible(var_name)
        #assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
        cat("Outliers successfully removed", "\n")
        return(invisible(dt))
    } else {
        cat("Nothing changed", "\n")
        return(invisible(df))
    }
}

```

```{r Load 2015 data, include=FALSE}
loadPkg('dplyr')
loadPkg('readxl')
loadPkg('readr')

#Read in general census data
census_2015 <- read.csv(file = 'acs2015updated.csv')

#Read in Specifically Education Data
edu_data_2015 <- read_csv("ACS_15_5YR_B15003_renamed.csv", skip = 1) %>% mutate(Id2 = as.numeric(Id2))

#Read in Freedom Data
freedom <- read_excel("Freedom_In_The_50_States_2018.xlsx", sheet = "Overall")

#Combine
census_edu_2015 <- left_join(census_2015, edu_data_2015, by = c("CensusTract" = "Id2"))
full_2015 <- full_join(census_edu_2015, freedom %>% filter(Year == 2015), by = "State") 

#For the education data, we will need to divide the count by the total
full_2015_eduProp <- full_2015 %>% mutate(PropHighSchool = EstHighSchool/EstTotal,
                                          PropBachelors = EstBachelors/EstTotal,
                                          PropDoctorate = EstDoctorate/EstTotal)
full_2015_varOfInterest <- full_2015 %>%  select(IncomePerCap,Hispanic:Asian,Professional:Production, SelfEmployed, Unemployment, FiscalPolicy,EconomicFreedom, RegulatoryPolicy, PersonalFreedom, OverallFreedom)

```

# Introduction


# Dataset 

# Description of Variables

## "First Attempts" Geo Plots

```{r geo scatter} 
full_2015_2 <- subset(full_2015, INTPTLONG > -150 & INTPTLONG < -50 & INTPTLAT > 24 & INTPTLAT < 50) 


loadPkg("ggplot2")
ggplot(full_2015_2, aes( x= INTPTLONG, y = INTPTLAT, color = Hispanic)) + 
  geom_point(size=0.005) + scale_color_gradient(low="orange", high="red") +
  labs (title="Scatter Plot: Spanish Ethnicity in 2015",
        x=" Latitude",
        y = "Longitude")

ggplot(full_2015_eduProp, aes( x= Professional, y = IncomePerCap, color = State == "New York", size = TotalPop)) +
  geom_point() +
  labs (title="Scatter Plot: IncomePerCap and Work type: Professional",
        x="Professional (% of pop in census tract)",
        y = "IncomePerCap")
```

# Baseline EDA of Income and Population

## Population Histogram and QQ

```{r population, include=FALSE} 
loadPkg("ggplot2")

# NA's
dat0<-full_2015[!is.na(full_2015$IncomePerCap),]
sum(is.na(dat0$IncomePerCap))

#removed outliers 
dat1 <- outlierKD2(dat0, IncomePerCap, TRUE)
#Histogram of dependent variable
hist(dat1$TotalPop, 
        col = "#a8c2fb",
        main = "population Count",
        xlab = "US Dollars")

#ggplot
qqnorm(dat1$TotalPop, main="Q-Q plot of TotalPop")
qqline(dat1$TotalPop)

#Mean after removing NA's
format(mean(dat1$TotalPop, na.rm = TRUE))

#Summary after removing NA's
summary(dat1$TotalPop)
#standard deviation
sd(dat1$TotalPop, na.rm = TRUE)

```

## Income Histogram and QQ

```{r income including outliers, include=FALSE} 
#Histogram of dependent variable
hist(dat0$IncomePerCap, 
        col = "#a8c2fb",
        main = "Income per Capita",
        xlab = "US Dollars")

#ggplot
qqnorm(dat0$IncomePerCap, main="Q-Q plot of IncomePerCap")
qqline(dat0$IncomePerCap)

#Mean after removing NA's
format(mean(dat0$IncomePerCap, na.rm = TRUE))

#Summary after removing NA's
summary(dat0$IncomePerCap)
#standard deviation
sd(dat0$IncomePerCap, na.rm = TRUE)
```

```{r income no outliers} 
#Histogram of dependent variable
hist(dat1$IncomePerCap, 
        col = "#a8c2fb",
        main = "Income per Capita",
        xlab = "US Dollars")

#ggplot
qqnorm(dat1$IncomePerCap, main="Q-Q plot of IncomePerCap")
qqline(dat1$IncomePerCap)

#Mean after removing NA's
format(mean(dat1$IncomePerCap, na.rm = TRUE))

#Summary after removing NA's
summary(dat1$IncomePerCap)
#standard deviation
sd(dat1$IncomePerCap, na.rm = TRUE)
```

# Individual EDA of independent variables.


```{r Freedom scores} 
loadPkg("dvmisc")
#Group by Economic Freedom
GroupEF <- quant_groups(dat1$EconomicFreedom, 4)
str(GroupEF)

#Boxplot of IncomePerCap by Economic Freedom
plot(IncomePerCap ~ GroupEF, data=dat1, main="IncomePerCap and GroupEF", col=c("#ffd18b","#ffb76d","#ffa264", "#ff875f") )
summary(dat1$EconomicFreedom)

#Histogram of Economic Freedom
hist(dat1$EconomicFreedom, 
        col = "#ffd18b",
        main = "Economic Freedom")

#qqplot of Econoimc Freedom
qqnorm(dat1$EconomicFreedom, main="Q-Q plot of Economic Freedom")
qqline(dat1$EconomicFreedom)


#Group by PersonalFreedom
GroupPF <- quant_groups(dat1$PersonalFreedom, 4)
str(GroupPF)

#Boxplot of IncomePerCap by Personal Freedom
plot(IncomePerCap ~ GroupPF, data=dat1, main="IncomePerCap and GroupPF", col=c("#BFBFFF","#A3A3FF", "#7879FF", "#4949FF") )
summary(dat1$PersonalFreedom)

#Histogram of Personal Freedom
hist(dat1$PersonalFreedom, 
        col = "#A3A3FF",
        main = "Personal Freedom")

#qqplot of Personal Freedom
qqnorm(dat1$PersonalFreedom, main="Q-Q plot of Personal Freedom")
qqline(dat1$PersonalFreedom)

#Group by Regulatory Policy
GroupRP <- quant_groups(dat1$RegulatoryPolicy, 4)
str(GroupRP)

#Boxplot of IncomePerCap by Regulatory Policy
plot(IncomePerCap ~ GroupRP, data=dat1, main="IncomePerCap and GroupRP", col=c("#FFDF01","#FED901", "#FFCF00", "#FEC300") )
summary(dat1$RegulatoryPolicy)

#Histogram of Regulatory Policy
hist(dat1$RegulatoryPolicy, 
        col = "#FFDF01",
        main = "Regulatory Policy ")

#qqplot of Regulatory Policy
qqnorm(dat1$RegulatoryPolicy, main="Q-Q plot of Regulatory Policy")
qqline(dat1$RegulatoryPolicy)


#Group by Fiscal Policy
GroupFP <- quant_groups(dat1$FiscalPolicy, 4)
str(GroupFP)
#Boxplot of IncomePerCap by Fiscal Policy
plot(IncomePerCap ~ GroupFP, data=dat1, main="IncomePerCap and GroupFP", col=c("#7EFFD4","#70EBBA", "#64D8A7", "#58BD95") )
summary(dat1$FiscalPolicy)

#Histogram of Fiscal Policy
hist(dat1$FiscalPolicy, 
        col = "#7EFFD4",
        main = "Fiscal Policy")

#qqplot of Fiscal Policy
qqnorm(dat1$FiscalPolicy, main="Q-Q plot of Fiscal Policy")
qqline(dat1$FiscalPolicy)

#Group by Overall Freedom
GroupOF <- quant_groups(dat1$OverallFreedom, 4)
str(GroupOF)
#Boxplot of IncomePerCap by Overall Freedom
plot(IncomePerCap ~ GroupOF, data=dat1, main="IncomePerCap and GroupOF", col=c("#C0C6CB","#98A5C0", "#7688BB", "#536CB5") )
summary(dat1$OverallFreedom)

#Histogram of Overall Freedom
hist(dat1$OverallFreedom, 
        col = "#A3A3FF",
        main = "Overall Freedom")

#qqplot of Overall Freedom
qqnorm(dat1$OverallFreedom, main="Q-Q plot of Overall Freedom")
qqline(dat1$OverallFreedom)
```

```{r Work variations}
#Group by Unemployment
GroupUnemployment <- quant_groups(dat1$Unemployment, 4)
str(GroupUnemployment)
#Boxplot of IncomePerCap by Unemployment
plot(IncomePerCap ~ GroupUnemployment, data=dat1, main="IncomePerCap and GroupUnemployment", col=c("#FFF57B","#FFE469", "#FECC51", "#FCB033") )
summary(dat1$Unemployment)

#Histogram of Unemployment
hist(dat1$Unemployment, 
        col = "#A3A3FF",
       )

#qqplot of Unemployment Unemployment
qqnorm(dat1$Unemployment, main="Q-Q plot of Unemployment")
qqline(dat1$Unemployment)

#Group by Professional
GroupProfessional <- quant_groups(dat1$Professional, 4)
str(GroupProfessional)
#Boxplot of IncomePerCap by Professional
plot(IncomePerCap ~ GroupProfessional, data=dat1, main="IncomePerCap and GroupProfessional", col=c("#F3D8F2","#E6B2E4", "#D88DD5", "#CA68C7") )
summary(dat1$Professional)

#Histogram of Professional
hist(dat1$Professional, 
        col = "#D88DD5",
        main = "Professional")

#qqplot of Professional
qqnorm(dat1$Professional, main="Q-Q plot of Professional")
qqline(dat1$Professional)


#Group by Office
GroupOffice <- quant_groups(dat1$Office, 4)
str(GroupOffice)
#Boxplot of IncomePerCap by Office
plot(IncomePerCap ~ GroupOffice, data=dat1, main="IncomePerCap and GroupOffice", col=c("#D1FFD5","#B4FFB2", "#98FF98", "#79F58A") )
summary(dat1$Office)
#Histogram of Office
hist(dat1$Office, 
        col = "#FECC51",
        main = "Office")

#qqplot of Office
qqnorm(dat1$Office, main="Q-Q plot of Office")
qqline(dat1$Office)


#Group by Service
GroupService <- quant_groups(dat1$Service, 4)
str(GroupService)
#Boxplot of IncomePerCap by Service
plot(IncomePerCap ~ GroupService, data=dat1, main="IncomePerCap and GroupService", col=c("#E0BCBF","#D8ABB1", "#CF989F", "#C0838C") )
summary(dat1$Service)

#Histogram of Service
hist(dat1$Service, 
        col = "#79F58A",
        main = "Service")

#qqplot of Service
qqnorm(dat1$Service, main="Q-Q plot of Service")
qqline(dat1$Service)

#Group by Construction
GroupConstruction <- quant_groups(dat1$Construction, 4)
str(GroupConstruction)
#Boxplot of IncomePerCap by Construction
plot(IncomePerCap ~ GroupConstruction, data=dat1, main="IncomePerCap and GroupConstruction", col=c("#A9AB98","#949180", "#7D7968", "#5E594F") )
summary(dat1$Construction)
#Histogram of Construciton
hist(dat1$Construction, 
        col = "#C0838C",
        main = "Construction")

#qqplot of Construction
qqnorm(dat1$Construction, main="Q-Q plot of Construction")
qqline(dat1$Construction)

#Group by Production
GroupProduction <- quant_groups(dat1$Production, 4)
str(GroupProduction)
#Boxplot of IncomePerCap by Production
plot(IncomePerCap ~ GroupProduction, data=dat1, main="IncomePerCap and Production", col=c("#F3D8F2","#E6B2E4", "#D88DD5", "#CA68C7") )
summary(dat1$Production)
#Histogram of Producrtion
hist(dat1$Production, 
        col = "#E6B2E4",
        main = "Production")

#qqplot of Production
qqnorm(dat1$Production, main="Q-Q plot of Production")
qqline(dat1$Production)

#Group by Self-Employed
GroupProduction <- quant_groups(dat1$SelfEmployed, 4)
str(GroupProduction)
#Boxplot of IncomePerCap by Selfemployed
plot(IncomePerCap ~ GroupProduction, data=dat1, main="IncomePerCap and Selfemployed", col=c("#F3D8F2","#E6B2E4", "#D88DD5", "#CA68C7") )
summary(dat1$SelfEmployed)
#Histogram of Self Employed
hist(dat1$SelfEmployed, 
        col = "#E6B2E4",
        main = "Production")

#qqplot of Self employed
qqnorm(dat1$SelfEmployed, main="Q-Q plot of Selfemployed")
qqline(dat1$SelfEmployed)
```

```{r  Ethnic Variatons}
loadPkg("dvmisc")
#Group by Black
GroupBlack <- quant_groups(dat1$Black, 4)
str(GroupBlack)
#Boxplot of IncomePerCap by Black Population
plot(IncomePerCap ~ GroupBlack, data=dat1, main="IncomePerCap and GroupBlack", col=c("#d0dfff","#a8c2fb", "#86abf9", "#6893ee") )
summary(dat1$Black)

#Histogram of Black
hist(dat1$Black, 
        col = "#d0dfff",
        main = "Black")

#qqplot of Black
qqnorm(dat1$Black, main="Q-Q plot of Black")
qqline(dat1$Black)

#Group by Hispanic
GroupHispanic <- quant_groups(dat1$Hispanic, 4)
str(GroupHispanic)
#Boxplot of IncomePerCap by Hispanic Population
plot(IncomePerCap ~ GroupHispanic, data=dat1, main="IncomePerCap and GroupHispanic", col=c("#F6BDC0","#F1959B", "#F07470", "#EA4C46") )
summary(dat1$Hispanic)
#Histogram of Hispanic
hist(dat1$Hispanic, 
        col = "#EA4C46",
        main = "Hispanic")

#qqplot of Hispanic
qqnorm(dat1$Hispanic, main="Q-Q plot of Hspanic")
qqline(dat1$Hispanic)

#Group by Asians
GroupAsian <- quant_groups(dat1$Asian, 4)
str(GroupAsian)
#Boxplot of IncomePerCap by Asian Population
plot(IncomePerCap ~ GroupAsian, data=dat1, main="IncomePerCap and GroupAsian", col=c("#FFE6C8","#FFCCBE", "#EFABA0", "#D6806F") )
summary(dat1$Asian)
#Histogram of Asian
hist(dat1$Asian, 
        col = "#FFE6C8",
        main = "Asian")

#qqplot of Asian
qqnorm(dat1$Asian, main="Q-Q plot of Asian")
qqline(dat1$Asian)

#Group by White
GroupWhite <- quant_groups(dat1$White, 4)
str(GroupWhite)
#Boxplot of IncomePerCap by White Population
plot(IncomePerCap ~ GroupWhite, data=dat1, main="IncomePerCap and GroupWhite", col=c("#F5F8FA", "#EFF2F4", "#EAEDEF", "#E0E3E5") )
summary(dat1$White)
#Histogram of White
hist(dat1$White, 
        col = "#F5F8FA",
        main = "White")

#qqplot of White
qqnorm(dat1$White, main="Q-Q plot of White")
qqline(dat1$White)

#Group by native
#GroupNative <- quant_groups(dat1$Native, 4)
#str(GroupNative)

#Boxplot of IncomePerCap by Native Population
#plot(IncomePerCap ~ GroupNative, data=dat1, main="IncomePerCap and GroupNative", col=c("#F3D8F2","#E6B2E4", "#D88DD5", "#CA68C7") )
summary(dat1$Native)
#Histogram of Native
hist(dat1$Native, 
        col = "#F5F8FA",
        main = "Native")

#qqplot of Native
qqnorm(dat1$Native, main="Q-Q plot of Native")
qqline(dat1$Native)


```

# Correlations

### !! full2015_varofInterest
```{r DF and cor Setup} 
dat1_varOfInterest <- dat1 %>% select(colnames(full_2015_varOfInterest))

full_2015_cor <- cor(dat1_varOfInterest, use = "complete.obs")

loadPkg("corrplot")
corrplot(full_2015_cor)
```

```{r Subsetting, include=FALSE} 
Freedom_2015 <- subset(dat1_varOfInterest, select = c(IncomePerCap, OverallFreedom, EconomicFreedom, PersonalFreedom, RegulatoryPolicy, FiscalPolicy))

Work_2015 <- subset(dat1_varOfInterest, select = c(IncomePerCap, Professional, Service, Office, Construction, Production, SelfEmployed, Unemployment))

Ethnic_2015 <- subset(dat1_varOfInterest, select = c(IncomePerCap, Hispanic, White, Black, Native, Asian))
```

```{r Cor w/i Freedom}
Freedom_2015_cor <- cor(Freedom_2015, use = "complete.obs")
#corrplot(Freedom_2015_cor)
corrplot.mixed(Freedom_2015_cor, tl.pos = "lt")
```


```{r Cor w/i Work}
Work_2015_cor <- cor(Work_2015, use = "complete.obs")
#corrplot(Work_2015_cor)
corrplot.mixed(Work_2015_cor, tl.pos = "lt")
```

```{r Cor w/i Ethnic Background}
Ethnic_2015_cor <- cor(Ethnic_2015, use = "complete.obs")
#corrplot(Ethnic_2015_cor)
corrplot.mixed(Ethnic_2015_cor, tl.pos = "lt")
```

# Scatter PLots

```{r}
ggplot(dat1, aes( x= Professional, y = IncomePerCap, color = White)) + 
  geom_point(size=0.1) + geom_smooth(method='lm', formula= IncomerPerCap~Professional) +
  labs (title="Scatter Plot of IncomePerCap and Professional",
        x="Professional",
        y = "IncomePerCap")

ggplot(dat1, aes( x= Unemployment, y = IncomePerCap, color = White)) + 
  geom_point(size=0.1) + geom_smooth(method='lm', formula= IncomerPerCap~Unemployment) +
  labs (title="Scatter Plot of IncomePerCap and Unemployment",
        x="Unemployment",
        y = "IncomePerCap")


ggplot(dat1, aes( x= Service, y = IncomePerCap, color = White)) + 
  geom_point(size=0.1) + geom_smooth(method='lm', formula= IncomerPerCap~Service) +
  labs (title="Scatter Plot of IncomePerCap and Service",
        x="Service",
        y = "IncomePerCap")


ggplot(dat1, aes( x= EconomicFreedom, y = IncomePerCap, color = White)) + 
  geom_point(size=0.1) + geom_smooth(method='lm', formula= IncomerPerCap~EconomicFreedom) +
  labs (title="Scatter Plot of IncomePerCap and EconomicFreedom",
        x="EconomicFreedom",
        y = "IncomePerCap")
```


# ANOVA

### !! full_2015_varOfInterest

```{r Majority Ethnicities Dataset} 
#' @title Returns the max value of each row of a data.frame or matrix
#'
#' @description
#' Returns maximum value of each row of a data.frame or matrix.
#' @param df Data.frame or matrix, required.
#' @param na.rm Logical value, optional, TRUE by default. Defines whether NA values should be removed first. Otherwise result will be NA when any NA is in the given vector.
#' @return Returns a vector of numbers of length equal to number of rows in df.
#' @template maxmin
#' @export
rowMaxs <- function(df, na.rm=TRUE) {

  if (is.matrix(df)) {df <- data.frame(df, stringsAsFactors=FALSE, drop = FALSE)}

  valid.cols <- sapply(df, function(x) { is.numeric(x) || is.logical(x) || is.character(x)})
  stopifnot(any(valid.cols))
  # or could just return NA?:
  # if (!any(valid.cols)) {return(NA)}
  if (any(!valid.cols) ) {warning('using only numeric (double or integer) or logical or character columns -- ignoring other columns ')}

  result <- do.call(pmax, c(df[ , valid.cols, drop = FALSE], na.rm=na.rm))

  result[nononmissing <- rowSums(!is.na(df[ , valid.cols, drop = FALSE]))==0] <- -Inf
  if (any(nononmissing)) {warning('where no non-missing arguments, returning -Inf')}
  return(result)

  # df = data.frame of numeric values, i.e. a list of vectors passed to pmax
  # Value returned is vector, each element is max of a row of df
}

dat1_varOfInterest1 <- dat1_varOfInterest %>% mutate(EthnicMajority = case_when(
  White >= 50 ~ "White",
  Hispanic >= 50 ~ "Hispanic",
  Black >= 50 ~ "Black",
  Native >= 50 ~ "Native",
  Asian >= 50 ~ "Asian",
  TRUE ~ "Diverse"))
EthnicityMaxes <- rowMaxs(dat1_varOfInterest1 %>%  select(White, Hispanic, Black, Native, Asian))
WorkMaxes <- rowMaxs(dat1_varOfInterest1 %>%  select(Professional:Unemployment))
dat1_varOfInterest2 <- cbind(dat1_varOfInterest1, EthnicityMaxes, WorkMaxes)

dat1_varOfInterest_withMajorityEthnicity <- dat1_varOfInterest2 %>% mutate(EthnicPlurality = case_when(
  White == EthnicityMaxes ~ "White",
  Hispanic == EthnicityMaxes ~ "Hispanic",
  Black == EthnicityMaxes ~ "Black",
  Native == EthnicityMaxes ~ "Native",
  Asian == EthnicityMaxes ~ "Asian",
  TRUE ~ "Error"),
  WorkPlurality = case_when(
  Professional == WorkMaxes ~ "Professional",
  Service == WorkMaxes ~ "Service",
  Office == WorkMaxes ~ "Office",
  Construction == WorkMaxes ~ "Construction",
  Production == WorkMaxes ~ "Production",
  SelfEmployed == WorkMaxes ~ "SelfEmployed",
  Unemployment == WorkMaxes ~ "Unemployment",
  TRUE ~ "Error"))
dat1_varOfInterest_withMajorityEthnicity$EthnicMajority <-
  as.factor(dat1_varOfInterest_withMajorityEthnicity$EthnicMajority)
dat1_varOfInterest_withMajorityEthnicity$EthnicPlurality <-
  as.factor(dat1_varOfInterest_withMajorityEthnicity$EthnicPlurality)
dat1_varOfInterest_withMajorityEthnicity$WorkPlurality <-
  as.factor(dat1_varOfInterest_withMajorityEthnicity$WorkPlurality)
```

```{r dataset} 
anova_dat <- dat1_varOfInterest_withMajorityEthnicity %>% filter(EthnicPlurality!= "Error",
                                                                 WorkPlurality!= "Error")
```

```{r Ethnic ANOVA}
Ethnic_2015_anova = aov(IncomePerCap ~ EthnicPlurality, data = anova_dat)
Ethnic_2015_anova
names(Ethnic_2015_anova)
summary(Ethnic_2015_anova)
```

```{r Ethnic Graph} 
ggplot(anova_dat, aes(x=EthnicPlurality, y=IncomePerCap)) + 
  geom_boxplot(outlier.shape=8, outlier.size=4) +
  labs(title="Income/Capita with Different Ethnic Majority", x="Ethnic Majority", y = "Income per Cap")
```

```{r Ethnic Tukey, include=FALSE}
TukeyHSD(Ethnic_2015_anova)
```

```{r Work ANOVA} 
Work_2015_anova = aov(IncomePerCap ~ WorkPlurality, data = anova_dat)
Work_2015_anova
names(Work_2015_anova)
summary(Work_2015_anova)
```

```{r Work Graph}
ggplot(anova_dat, aes(x=WorkPlurality, y=IncomePerCap)) + 
  geom_boxplot(outlier.shape=8, outlier.size=4) +
  labs(title="Income/Capita with Different Work Majority", x="Work Majority", y = "Income per Cap")
```

```{r Work Tukey, include=FALSE}
TukeyHSD(Work_2015_anova)
```

# Chi Squared Tests

```{r Chi Squared Test} 
loadPkg('sjPlot')
loadPkg('visualize')

chi_table_dat <- dat1_varOfInterest_withMajorityEthnicity %>%
  select(c(WorkPlurality, EthnicPlurality)) %>% filter(WorkPlurality != "Error")

chi_table_dat %>%
  sjtab(fun = "xtab", var.labels=c("Work", "Ethnicity"),
       show.summary=T, show.exp=T, show.legend=T)
```

# Regression
##Exhaustive Search

```{r} 
loadPkg('leaps')
#This is essentially best fit 
reg.best10 <- regsubsets(IncomePerCap~. , data = dat1_varOfInterest, nvmax = 17)  
plot(reg.best10, scale = "adjr2", main = "Adjusted R^2")
plot(reg.best10, scale = "r2", main = "R^2")
# In the "leaps" package, we can use scale=c("bic","Cp","adjr2","r2")
plot(reg.best10, scale = "bic", main = "BIC")
plot(reg.best10, scale = "Cp", main = "Cp")
reg.summary<-summary(reg.best10)
names(reg.best10)
plot(reg.summary$rsq,xlab = "# of Varibales", ylab = "Rsquare", type = "l")
plot(reg.summary$rss,xlab = "# of Varibales", ylab = "RSS", type = "l")
#Adjusted R2
which.max(reg.summary$adjr2)
plot(reg.summary$adjr2,xlab = '# of Varibales', ylab = 'Adjusted Rsq', type = "l")
points(12,reg.summary$adjr2[12],col = "red",cex = 2, pch =20)
#CP
which.min(reg.summary$cp)
plot(reg.summary$cp,xlab = '# of Varibales', ylab = 'CP', type = "l")
points(12,reg.summary$cp[12],col = "red",cex = 2, pch =20)
#BIC
which.min(reg.summary$bic)
plot(reg.summary$bic,xlab = '# of Varibales', ylab = 'BIC', type = "l")
points(10,reg.summary$bic[12],col = "red",cex = 2, pch =20)

```

## Forward Selection

```{r}
reg.forward10 <- regsubsets(IncomePerCap~. , data = dat1_varOfInterest, nvmax = 17, nbest = 2, method = "forward")
plot(reg.forward10, scale = "adjr2", main = "Adjusted R^2")
plot(reg.forward10, scale = "bic", main = "BIC")
plot(reg.forward10, scale = "Cp", main = "Cp")
#summary(reg.forward10)
```


## Backward Selection

Now backwards (`nvmax=10` and `nbest=2`)

```{r}
reg.back10 <- regsubsets(IncomePerCap~. , data = dat1_varOfInterest, nvmax = 17, nbest = 2, method = "backward")
plot(reg.back10, scale = "adjr2", main = "Adjusted R^2")
plot(reg.back10, scale = "bic", main = "BIC")
plot(reg.back10, scale = "Cp", main = "Cp")
#summary(reg.back10)
```

## Sequential Replacement seqrep  

```{r}
reg.seqrep <- regsubsets(IncomePerCap~. , data = dat1_varOfInterest, nvmax = 17, nbest = 2 , method = "seqrep")
plot(reg.seqrep, scale = "adjr2", main = "Adjusted R^2")
plot(reg.seqrep, scale = "bic", main = "BIC")
plot(reg.seqrep, scale = "Cp", main = "Cp")
#summary(reg.seqrep)
```

# Conclusion